Recruitment rate estimation in waterfowl using a non-invasive method

### 2021-12-09 Code/Draft_ruddy_duck_recruitment_rate.Rmd
Discussion update

---
lang: en-EN
bibliography: C:/Users/Adrien/Dropbox/JabRef/Library_OFB.bib
csl: https://www.zotero.org/styles/apa
link-citations: true
title: "Recruitment rate estimation in waterfowl using a non-invasive method"
author: 
- Adrien Tableau^[Office Français de la Biodiversité, France - adrien.tableau@ofb.gouv.fr]
# - Alain Caizergues^[Office Français de la Biodiversité, France - alain.caizergues@ofb.gouv.fr]
- Iain Henderson^[Animal and Plant Health Agency, United Kingdom - iain.henderson@apha.gov.uk]
- Sébastien Reeber^[Société Nationale pour la Protection de la Nature, France - sebastien.reeber@snpn.fr]
# - Matthieu Guillemain^[Office Français de la Biodiversité, France -  matthieu.guillemain@ofb.gouv.fr]
date: "`r format(Sys.time(), '%d %B, %Y')`"
abstract: |
  The response of a waterfowl population to a harvest pressure depends on its capacity to renew. The recruitment, i.e. the number of young adults reproducing for the first time, is a key indicator to describe the renewal of a population, and therefore an essential tool in population management. The recruitment rate, defined inhere as the number of recruits produced per breeder, is even more informative because it is independent of the breeding population size and allows comparisons over time and between species. The proportion of young adult in a waterfowl population, which is a main step in estimating the recruitment rate, is often estimated from game-hunting samples. However, this proportion is not accessible in years without harvest, or with a low harvest rate. Moreover, the age-structure in the harvest samples does not necessarily reflect the underlying age-structure of the population. It is often skewed towards immatures and can lead to an overestimation of the recruitment rate. In waterfowl, adult males usually display brighter colours than females and immatures from both sexes. This dichromatism can be characterized and monitored from count surveys, a non-invasive method. Such information can be used to infer the proportion of immatures, and consequently the recruitment rate. In using two populations of ruddy duck, *Oxyura oxyura*, this study introduces a model to estimate the recruitment rate from count data. To test the accuracy of the method, the results are compared to recruitment rate estimates from samples, and their consistency to support the observed population growth rates is also tested. The results suggest that the counting method is a versatile tool to estimate the recruitment rate if the species is monitored during the appropriate time window. Authors argue for considering a two-category count protocol in winter surveys of waterfowl species for which adult males can be differentiated from other individuals.

  **Keywords**: fecundity - productivity - reproductive success - age ratio - juvenile - duck
output: 
    bookdown::pdf_document2:
      latex_engine: xelatex
      toc: false
    bookdown::word_document2:
      toc: false
header-includes:
  \usepackage{float}

---

@@

# Accuracy of the counting method

As expected, countalways lower than samples. Variability consistent to the samples as well, but slightly enhanced compared to the raw immature proportion \@ref(fig:sup)

@@

A significant change in adult sex ratio would bias the estimation of recruitment rate. Since this measure integrates many bird cohorts, especially if it is a long-living species, updates on this ratio are not necessary on a year based

#

# Impact of harvest on recruitment rate

## reparler des notions en introduction

figure 8 -> One said that recruitment rate can increase if harvest because of heterogeneity and density dependence, but it is not what we observe in France, likely because the exploitation is high inhere and consequently impact.

obviously, harvest impact the pre

rajouter en intro cette notion.
a same value of age ratio can correspond to different fecundity levels, so warning. 
a same value of age ratio for a population in depletion and a population stabilized are not similary productive, the first one is less productive than the second one ! 
so warning, focus on recruitment rate and not on age ratio if the dynamics of the population varies! recruitment proportion is commonly used in to discuss the fecundity of a population, but unlike the recruitment rate, this indicator does not consider the dynamics of the population and does reflect fecundity only on stable growth period

@@

@@

# Aknowledgements {-}

This work was partly funded by the LIFE Oxyura project (LIFE17 NAT/FR/000942) through the LIFE program. This work was carried on with the impulse of Jean-François Maillard, and Jean-Baptiste Mouronval from the *Office Français de la Biodiversité*, and Jean-Marc Gillier from the *Société Nationale pour la Protection de la Nature*. The authors aknowledge all the contributors of the data collection in the UK and in France, especially: **UK names**, Vincent Fontaine, Denis Lacourpaille, Jules Joly, Justin Potier, Valentin Boniface, Alexis Laroche, and Médéric Lortion, **other names**.

>>>>

---
lang: en-EN
bibliography: C:/Users/adrien.tableau/Dropbox/JabRef/Library_OFB.bib
csl: https://www.zotero.org/styles/apa
link-citations: true
title: "Recruitment rate estimation in waterfowl using a non-invasive method"
author: 
- Adrien Tableau^[Office Français de la Biodiversité, France - adrien.tableau@ofb.gouv.fr]
# - Alain Caizergues^[Office Français de la Biodiversité, France - alain.caizergues@ofb.gouv.fr]
- Iain Henderson^[Animal and Plant Health Agency, United Kingdom - iain.henderson@apha.gov.uk]
- Sébastien Reeber^[Société Nationale pour la Protection de la Nature, France - sebastien.reeber@snpn.fr]
# - Matthieu Guillemain^[Office Français de la Biodiversité, France -  matthieu.guillemain@ofb.gouv.fr]
date: "`r format(Sys.time(), '%d %B, %Y')`"
abstract: |
  The response of a waterfowl population to a harvest pressure depends on its capacity to renew. The recruitment, i.e. the number of young adults reproducing for the first time, is a key indicator to describe the renewal of a population, and therefore an essential tool in population management. The recruitment rate, defined inhere as the number of recruits produced per breeder, is even more informative because it is independent of the breeding population size and allows comparisons over time and between species. The proportion of young adult in a waterfowl population, which is a main step in estimating the recruitment rate, is often estimated from game-hunting samples. However, this proportion is not accessible in years without harvest, or with a low harvest rate. Moreover, the age-structure in the harvest samples does not necessarily reflect the underlying age-structure of the population. It is often skewed towards immatures and can lead to an overestimation of the recruitment rate. In waterfowl, adult males usually display brighter colours than females and immatures from both sexes. This dichromatism can be characterized and monitored from count surveys, a non-invasive method. Such information can be used to infer the proportion of immatures, and consequently the recruitment rate. In using two populations of ruddy duck, *Oxyura oxyura*, this study introduces a model to estimate the recruitment rate from count data. To test the accuracy of the method, the results are compared to recruitment rate estimates from samples, and their consistency to support the observed population growth rates is also tested. The results suggest that the counting method is a versatile tool to estimate the recruitment rate if the species is monitored during the appropriate time window. Authors argue for considering a two-category count protocol in winter surveys of waterfowl species for which adult males can be differentiated from other individuals.

  **Keywords**: fecundity - productivity - reproductive success - age ratio - juvenile - duck
output: 
    bookdown::pdf_document2:
      latex_engine: xelatex
      toc: false
    bookdown::word_document2:
      toc: false
header-includes:
  \usepackage{float}
---

@@

## Accuracy of the counting method

*comparison to sample estimation*

As expected, count always lower than samples. Variability consistent to the samples as well, but slightly enhanced compared to the raw immature proportion \@ref(fig:sup)

@@

A significant change in adult sex ratio would bias the estimation of recruitment rate. However, since this ratio integrates many age-cohorts, the temporal autocorrelation is structurally strong. Significant variations can then occur only every multiple years, especially if it is a long-living species. Consequently, it is not necessary to monitor and update the adult sex ratio on a yearly base.

*survival consistency*

survival in the range of the expected values -> so no apparent issue of scaling issue, 
so we the hypothesis that male-like individuals corresponds only to adult male is validated.

*average recruitment rate compared to a proxy of maximum recruitment rate*

always below the maximum recruitment rate so satisfying.

overall, even if we don't have access to the true value of the recruitment rate to test the accuracy of the counting method, a beam of arguments support that the method is not heavily biased, unlike the sampling method, as demonstrated in @Fox2014, even if one considers a picture limited to the end of the hunting season [@Fox2016].

## Impact of harvest

Survival higher than one but warning, observation error, so annual population growth rate might be inaccurate on some years. Also, we assumed a closed population, but it is possible that years with adult survival higher than one corresponds to an arrival of birds from the other population. As a consequence, the recruitment rate on these years may also be overestimated.

No impact on the adult survival in France, but it seems that there is an impact in UK.
Why, because harvest occurs mostly in summer in France, and in winter in UK.

Impact on the recruitment rate in France, and no impact in UK.

Overall, the harvest seems to impact differently France and UK, likely because the harvest mortality occurs mostly in summer in France, and in winter in UK.

But caution, a population model linking the harvest rate to state variables describing the population is necessary to check precisely the impact of the harvest pressure, and test hypothesis on the functioning of harvest mortality.

Recruitment rate decrease with the harvest for the French population, and the UK population recruitment rate is also below the French one. 
it seems quite clear that the harvest affects the recruitment rate.

*reference to introduction*

It is quite consistent with the assumption in introduction. Since studied populations are in colonization of their habitat, maximum densities are not reached, and then no or limited density dependent effect occur.

@@

it may be possible that first sight of density dependence occur in UK from 1000 individuals, but was not big enough to fully curb the population growth. Since there is a strong correlation between the number of birds observed and the number of sites occupied by the birds, this inflexion may be simply due to a limited number of new favourable sites to be colonized.

What about heterogeneity, figure 8 -> One said that recruitment rate can increase if harvest because of heterogeneity and density dependence, but it is not what we observe in France, likely because the exploitation is high inhere and consequently impact.

## Use of recruitment rate

*better than sex ratio*
a same value of age ratio can correspond to different fecundity levels, so warning. 
a same value of age ratio for a population in depletion and a population stabilized are not similary productive, the first one is less productive than the second one ! 
so warning, focus on recruitment rate and not on age ratio if the dynamics of the population varies! recruitment proportion is commonly used in to discuss the fecundity of a population, but unlike the recruitment rate, this indicator does not consider the dynamics of the population and does reflect fecundity only on stable growth period

@@

# Acknowledgments {-}

This work was partly funded by the LIFE Oxyura project (LIFE17 NAT/FR/000942) through the LIFE program. This work was carried on with the impulse of Jean-François Maillard, and Jean-Baptiste Mouronval from the *Office Français de la Biodiversité*, and Jean-Marc Gillier from the *Société Nationale pour la Protection de la Nature*. The authors acknowledge all the contributors of the data collection in the UK and in France, especially: **UK names**, Vincent Fontaine, Denis Lacourpaille, Justin Potier, Alexis Laroche, Médéric Lortion, Jules Joly, and Valentin Boniface.
